Data Specialization for a Staged Lambda Calculus

### 2015-02-27 paper/intro.tex
pass on the abstract

While partial evaluation and metaprogramming have been studied extensively,
frequency reduction techniques have not seen the same attention. In
\cite{JS86-staging}, J{\o}rring and Scherlis proposed {\em pass separation} as a
manual method for {\em splitting} a program into multiple passes.
%
In our example, pass separation would split the function $f$ into two passes
$f_1$ and $f_2$ such that $f(x,y_i) = f_2(f_1(x),y_i)$. Then we evaluate $F$ by
evaluating the first pass on $x$, and then the second pass on each $y_i$.
%
\begin{lstlisting}
fun @$F_\textrm{staged}(x, y_0, y_1, \ldots, y_{m-1})$@ = 
  let @$fx = f_1(x)$@
  in @$(f_2(fx, y_0);  f_2(fx, y_1)  \ldots f_2(fx, y_{m-1}))$@.
\end{lstlisting}
%
The key difference between partial evaluation (or metaprogramming) and pass
separation is that the latter can be performed without access to the first
argument $x$: $F_\textrm{staged}$ works for any $x$, while $F_v$ is defined
only for $x=v$. (Of course, when a particular $v$ \emph{is} known, one may
always use partial evaluation to specialize $F_\textrm{staged}$ to that $v$.)

>>>>

While partial evaluation and metaprogramming have been studied extensively,
frequency reduction techniques have not seen the same attention. In
\cite{JS86-staging}, J{\o}rring and Scherlis proposed {\em pass separation} as a
manual method for {\em splitting} a program into multiple passes.
%
In our example, pass separation would split the function $f$ into two passes
$f_1$ and $f_2$ such that $f(x,y_i) = f_2(f_1(x),y_i)$. Then we evaluate $F$ by
evaluating the first pass on $x$, and then the second pass on each $y_i$.
%
\begin{lstlisting}
fun @$F_\textrm{multipass}(x, y_0, y_1, \ldots, y_{m-1})$@ = 
  let @$fx = f_1(x)$@
  in @$(f_2(fx, y_0);  f_2(fx, y_1)  \ldots f_2(fx, y_{m-1}))$@.
\end{lstlisting}
%
The key difference between partial evaluation (or metaprogramming) and
pass separation is that the latter can be performed without access to
the first argument $x$: $F_\textrm{multipass}$ works for any $x$,
while $F_v$ is defined only for $x=v$.  Note that unlike
precomputation, frequency reduction does not require knowledge of
arguments; if a particular $v$ is known, we can use partial evaluation
to specialize $F_\textrm{multipass}$ to that $v$.
